Using reference models in feature selection problems \vspace{.1in}

### 2019-07-24 paper/ref_approach.tex
clean and break long lines for easier diffing

%In statistical applications, one of the main steps of the modelling workflow is covariate or feature selection, which can be seen as a special case of model reduction. From a full Bayesian perspective, when the goal of the analysis is to do prediction, no selection should be carried out and integration over all parameters uncertainties, i.e. posterior distributions, is the recommended procedure. In a high dimensional parameter space the choice of the prior plays an even more key role due to different difficulties that can rise, as computation burden and overfitting. In addition to that, sometimes collecting all the features for a future observation is not possible or too expensive. In order to solve or, at least, simplify these problems, different feature selection procedures have been proposed both in the frequentist and the Bayesian setting \cite[e.g. see][]{paper:feature_selection,vehtari2012survey,paper:model_selection}. Sometimes the task of the analysis is not prediction, but rather inference on which variables are substantially related to the phenomenon of interest. This typically happens in a \textit{small n large p} scenario, as in many microarray data analysis in medicine or biology. A very large number of genes is usually analysed in two groups of individuals labeled as \textit{positive} or \textit{negative} to some disease. The main goal of the analysis is to spot the subset of explanatory genes in order to perform further experiments \cite[see examples in][]{efron2012large}.

% The necessity of carrying out the selection arises from different problematics. Sometimes the leading objective of the analysis is to predict future observationgs, thus the main focus in building the model is on the predictive performance. Yet a high dimensional parameter space often leads to dangers as overfitting or other technical problems, as expensive costs to collect all the features for future observations or computational and memory burdens (e.g. when operating on limited resources), and thus a sparser representation of the model would be preferred. In addition to that, simpler models bring also advantages in terms of interpretability. Some other times,
% variable selection can be also seen as an inferential tool: the interest is on knowing which variables are meaningful, or relevant, in relation to the target variable of interest.
% This typically happens in a \textit{small n large p} scenario, as in many microarray data analysis in medicine or biology.

% A very large number of genes is usually analysed in two groups of individuals labeled as \textit{positive} or \textit{negative} to some disease; the goal is to spot the subset of explanatory genes in order to perform further experiments \cite[see examples in][]{efron2012large}.

% process is going to be based on assumptions regarding the ``true''
% data generation mechanism.  Making assumptions means making models,
% thus we identify two families of feature selection, or more in general
% model reduction, approaches depending on how the data generation
% mechanism is modelled: the data based and the reference model based
% ones.

% The reference model approach has been used for long time in the Bayesian framework, an example the pioneering work by \cite{paper:reference_lindley}. Following \cite{goutis1998model}, \cite{paper:original_proj} project the posterior distribution of the reference model in each candidate model by minimisation of the Kullback-Liebler divergence, and then select the smallest one with enough explanatory power avoiding any prior elicitation for the submodels. In the special case of candidate models being GLMs from the exponential family, the optimisation problem corresponds to the task of finding the maximum likelihood estimate using the expected pointwise predictions of the reference model instead of the observed target values. Different extensions have followed \cite{paper:original_proj}, as \cite{nott2010bayesian} that add an $l_{1}$ penalty to select the best approximating model and \cite{tran2012predictive}, who select the best approximating model by optimising the Kullback-Liebler divergence between the predictive distribution of the reference model and the approximating model plus an $l_{1}$ penalty on the submodel coefficients. A recent review of these methods together with some methodological improvements, as clustered projection and selection by comparison of a predictive performance utility, is given by \cite{paper:projpred}.

% and shows some examples of how it can be used to seek a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model.

% \cite{faraggi2001understanding} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups.

%\cite{harrell2015regression} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
%\cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance of the explored submodels with the reference model one.
%\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and, when it is used, on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

% We hope that this work can increase
% the attractiveness on this family of methods and, therefore, it can be
% of interest to a large number of practitioners.

@@

% In Figure \ref{fig:correlation} the reference model filters the noise.
% This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}.\\
% Above the correlation was used to estimate the relevance of a covariate.  
As an another example, let us consider a candidate submodel in a feature selection process and let us refer to the parameter distribution of the submodel by $\pi$ and to the induced predictive distribution by $q_{\pi}(\tilde{y})$. We would like to choose $\pi$ so that $q_{\pi}(\tilde{y})$ maximises some predictive performance utility, for example, the expected log-predictive density (elpd) defined as:
\
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y},
\end{equation}
where $p_{t}(\tilde{y})$ denotes the (usually unknown) true generating mechanism of future data $\tilde{y}$. If we refer to the posterior predictive distribution of a reference model with $p(\tilde{y}|D)$, where $D$ stands for the data on which we conditioned on, we can approximate \eqref{eq:elpd} using $p(\tilde{y}|D)$ instead of the true data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the elpd using the reference model predictive distribution is equivalent to the minimisation of the Kullback-Leibler (KL) divergence from the reference model's predictive distribution to the submodel's predictive distribution:
\
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter} describes what is referred to as the projection of the predictive distribution, which is the general idea behind the projection predictive approach \cite[see][]{paper:projpred}. Again the reference model is acting as a filter, or substitute, on data.

%--- OR ---

%What we do in Figure \ref{fig:correlation} is to use the reference model predictive ability as a noise-filter on the observed data. This simple idea is what happens every time a reference model is used, as it is evident for example in the preconditioning by \cite{paul2008preconditioning} or in the draw-by-draw projection for GLMs by \cite{paper:original_proj}. Also the general idea of the projection of the predictive can be seen under this perspective. Indeed, let us call $p_{\text{ref}}(\tilde{y}|D)$ the posterior predictive distribution for a new observation $\tilde{y}$ for the reference model and let $\pi(\theta)$ and $q_{\pi}(\tilde{y})$ be respectively the submodel parameter and the induced predictive distributions. The projection of the predictive distribution allows to find the optimal parameter distribution $\pi^{*}$ minimising the KL discrepancy among the induced predictive distribution of the submodel and the reference model one:
%\
%\begin{equation} 
%\pi^{*}\quad \text{s.t.} \quad \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi^{*}}(\tilde{y})]=\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
%\end{equation}
%Consider now the predictive performance of such submodel, a common utility in Bayesian context is the expected log-predictive density defined as:
%\
%\begin{equation} 
%\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y} 
%\end{equation}
%where $p_{t}$ is the distribution of a new observation under the true data generation mechanism. Maximising the elpd of a model is equivalent to minimising the KL-divergence of its predictive distribution from the true data generation mechanism. Therefore the projection corresponds to maximising the elpd using as distribution of the true data generation mechanism the posterior predictive distribution of the reference model, i.e.:
%\
%\begin{equation} \label{proj_as_filter}
%\underset{\pi}{\text{min}} \; \text{KL}[p_{\text{ref}}(\tilde{y}|D)||q_{\pi}(\tilde{y})] \leftrightarrow \underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p_{\text{ref}}(\tilde{y}|D)d\tilde{y} 
%\end{equation}
%Equation \eqref{proj_as_filter} clearly shows the action of the reference model as a filter on our target observations, substituting the unknown true data generation mechanism, that is usually estimated with the data empirical distribution, with the best known approximation that is available, i.e. the reference model posterior predictive distribution.

%------------------

@@

%we compare some well-known methods from the literature using both such reference model approach and the original data, observing at the end of the selection higher stability and quality of the selection thanks to the reference model.

@@

As a motivational example to the implications of using a reference model in feature selection problems, we show its application in the form of the projection predictive approach, which uses the reference model to project the posterior draws to the submodels \citep{paper:original_proj}. The \texttt{R}-package \texttt{projpred} (available at \url{https://CRAN.R-project.org/package=projpred}) implements the projection in case of submodels in the family of the generalised linear models and, additionally, it provides a framework to choose the optimal submodel size through a predictive utility comparison between the submodels and the reference model. 
\\
We now summarise the workflow of the projection predictive approach in the particular case of the draw-by-draw projection \cite[original formulation by][]{paper:original_proj}, following \cite{paper:projpred}. Suppose we have observed $n$ statistical units with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed covariates for which we want to obtain a minimally relevant subset. Than, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of $S$ draws from the reference model posterior.
\item Rank the covariates according to their relevance using some heuristics and consider as candidate submodels only those which preserve this order,  starting from including only the highest ranked covariate (the submodels are then naturally identified by their model size). This is not strictly necessary but reduces the number of submodels to consider in the following steps.
\item For each submodel, project the reference model posterior draws as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL} \left( p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\;||\;q(\tilde{y}_{i}|\boldsymbol{\theta}^{s}) \right), \quad s=1,..,S \,,
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the predictive distribution of the reference model with parameters fixed at $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariate values related to the statistical unit (identified by the subscript $i$), whereas $q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ is the predictive distribution of the submodel. The projected draws $\boldsymbol{\theta}_{\perp}^{s}$ then present the projected posterior for the submodel.
\item For each submodel (size), test the predictive performance, for example, via cross-validation.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model's predictive utility score.
\end{enumerate}

@@

%\begin{table}[tp]
%\scriptsize
%\centering
%\begin{tabular}{l||l|r||l|r}
%  \hline
%M & projpred & Freq (\%) & steplm & Freq (\%)  \\ 
%  \hline
%1 & abdomen, weight & 34.2 & abdomen, height, wrist, age, chest, biceps & 3.2 \\
%2 & abdomen, wrist & 13.3 & abdomen, height, wrist, age, neck, forearm, thigh, hip & 2.9 \\
%3 & abdomen, weight, wrist & 7.6 & abdomen, height, wrist, age, forearm, chest & 1.9 \\
%4 & abdomen, height & 7.4 & abdomen, height, wrist, age, neck, forearm, chest & 1.9 \\
%5 & abdomen, height, wrist & 6.7 & abdomen, height, wrist, age, neck, forearm, chest, thigh, hip & 1.9 \\
%6 & abdomen, age, wrist & 2.9 & abdomen, height, wrist, age, neck, chest, biceps & 1.8 \\
%7 & abdomen, height, neck & 2.2 & abdomen, height, wrist, age, neck, thigh, biceps, hip & 1.6 \\
%8 & abdomen, age, height, wrist & 1.2 & abdomen, height, wrist, age, neck, forearm & 1.5 \\
%9 & abdomen, weight, thigh & 1.1 & abdomen, height, wrist, age, neck, biceps & 1.5 \\
%10 & abdomen, weight, neck & 1.0 & abdomen, height, wrist, age, beck, forearm, chest, biceps & 1.4 \\
%   \hline
%\end{tabular}
%\caption{Bootstrap model selection frequencies after 1000 bootstrap samples.}
%\label{tab:model_frequencies}
%\end{table}

@@

%\begin{table}[tp]
%\scriptsize
%\centering
%\begin{tabular}{l|r|r|r|r}
%  \hline
% & rmse.full & rmse.sel & size.sel & noisy.sel \\ 
%  \hline
%projpred & 4.4 & 4.5 & 2 & -  \\
%steplm & 4.4 & 4.5 & 7 & - \\
%\hline
%\hline
%projpred & $^{*}$4.4 & $^{*}$4.5 & 2 & 0  \\
%steplm & 5.3 & 5.5 & 38 & 32 \\
%   \hline
%\end{tabular}
%\caption{Model predictive performances with original data (first two rows) and adding noisy features (last two rows). Results with an asterisk using 10-fold CV, otherwise 20-fold CV.}
%\label{tab:model_performances}
%\end{table}

@@

\hypertarget{shrinkage-signal}{%
\subsubsection{Shrinkage and signal analysis}\label{shrinkage-signal}}
In the framework of fully Bayesian estimation methods, we analyse the effect of the reference model using two continuous hierarchical shrinkage priors: the Dirichlet-Laplace \citep{paper:dirichlet_laplace} and the regularised horseshoe prior \citep{paper:rhs}. 
% Although the horseshoe+ prior by \cite{bhadra2017horseshoe+} is a state-of-the-art prior for sparse signal analysis in the context of the normal means problem, we decided to include the regularised horseshoe due to its faster MCMC inference, since our purpose in this paper is not to study the best prior choice, but rather compare and show the benefits of using a reference model when carrying out variable selection.
For the Dirichlet hyperparameter of the Dirichlet-Laplace prior, we use an uniform prior on the interval $(1/p,1/2)$. For the regularised horseshoe prior, we follow the recommendations provided in \cite{paper:rhs} by setting the prior guess of the number of relevant covariates to one in order to achieve the highest possible shrinkage. The results are not sensitive with respect to the specific choice.
\
Since, as explained at the start of Section \ref{comparison}, continuous priors usually do not provide an automatic selection procedure, we study the amount of shrinkage and the bias of the posterior medians looking at the SESE and the SSE (see definitions \eqref{eq:SSE} and \eqref{eq:SESE}), together with the posterior mean of the shrinkage factors.

@@

%The maximum likelihood estimator is used to estimate the variance of the null hypothesis in the local false discovery rate \cite[see][Chap. 6]{efron2012large}, whereas the median absolute deviation from zero for the empirical Bayes median method. The selection via posterior credible intervals is done using a regularised horseshoe prior on the parameters for the means, while a standard log-normal  prior for the variance

@@

%\begin{figure}[tp]
%  \centering
%  \includegraphics[width=0.98\textwidth]{graphics/bodyfat_varyingP_sensitivity_vs_fdr.pdf}
%  \caption{Sensitivity against false discovery rate based on 100 data simulations.\\}
%  \label{fig:bodyfat_varyingP_sensitivity_vs_fdr}
%\end{figure}

%\begin{figure}[tp]
%  \centering
%  \includegraphics[width=0.98\textwidth]{graphics/bodyfat_varyingP_stability.pdf}
%  \caption{Stability estimates with 0.95 confidence intervals based on 100 data simulations.\\}
%  \label{fig:bodyfat_varyingP_stability}
%\end{figure}

% As an additional experiment, we run the selection bootstrapping from the whole dataset, but varying the overall number of covariates from $p=100$ up to $p=5000$. Figures \ref{fig:bodyfat_varyingP_sensitivity_vs_fdr} and \ref{fig:bodyfat_varyingP_stability} respectively show the sensitivity against the false discovery rate and the stability results. As in the previous experiment, we note that the reference model has a larger impact as the complexity of the problem increases: previously as $n$ decreases, now as $p$ increases. Note the different scale of the sensitivity and the false discovery rate axes. We observe an odd behaviour of the selection via credible intervals (ci.90): when $p=100$ and $p=5000$ the results are strongly different from the other methods, both in terms of sensitivity/false discovery rate and stability. 
%Such behaviour can be in part explained by the connection between low stability and high uncertainty in the estimates.

@@

%Also, in this case, we see that there is no notable difference in terms of the stability of the two approaches, which in this case is qualitatively assessed through the variability of the histogram. We can not use the stability measure provided by \cite{paper:stability}, since it does not take in account possibly swaps in the set of selected features due to high correlation between the covariates, which in this case can occur.

@@

%In a real-world data application (body fat data), we have applied one of these methods, namely the projection predictive approach as indicated by \cite{paper:projpred}, in opposition to the selection via stepwise backward regression with the results provided by \cite{paper:bodyfat}. The selection via projection resulted to be more stable and led to a sparser representation. After adding additional noisy covariates, the selected submodel showed also higher predictive performance compared to the one selected by the stepwise backward regression, which included several non-relevant features. Applying a reference model on top of the stepwise backward regression resulted in a minor number of non-relevant features selected and an improved predictive performance, yet without reaching the projection predictive approach results.

>>>>

@@

As an another example, let us consider a candidate submodel in a
feature selection process and let us refer to the parameter
distribution of the submodel by $\pi$ and to the induced predictive
distribution by $q_{\pi}(\tilde{y})$. We would like to choose $\pi$ so
that $q_{\pi}(\tilde{y})$ maximises some predictive performance
utility, for example, the expected log-predictive density (elpd)
defined as: \
\begin{equation}\label{eq:elpd}
\text{elpd}[q_{\pi}]=\int \text{log}\,q_{\pi}(\tilde{y})p_{t}(\tilde{y})d\tilde{y},
\end{equation}
where $p_{t}(\tilde{y})$ denotes the (usually unknown) true generating
mechanism of future data $\tilde{y}$. If we refer to the posterior
predictive distribution of a reference model with $p(\tilde{y}|D)$,
where $D$ stands for the data on which we conditioned on, we can
approximate \eqref{eq:elpd} using $p(\tilde{y}|D)$ instead of the true
data generation mechanism $p_{t}(\tilde{y})$. The maximisation of the
elpd using the reference model predictive distribution is equivalent
to the minimisation of the Kullback-Leibler (KL) divergence from the
reference model's predictive distribution to the submodel's predictive
distribution: \
\begin{equation} \label{proj_as_filter}
\underset{\pi}{\text{max}} \; \int \text{log}\,q_{\pi}(\tilde{y})p(\tilde{y}|D)d\tilde{y} \leftrightarrow \underset{\pi}{\text{min}} \; \text{KL}[p(\tilde{y}|D)||q_{\pi}(\tilde{y})] 
\end{equation}
The term on the right-hand side of Equation \eqref{proj_as_filter}
describes what is referred to as the projection of the predictive
distribution, which is the general idea behind the projection
predictive approach \cite[see][]{paper:projpred}. Again the reference
model is acting as a filter, or substitute, on data.

@@

@@

As a motivational example to the implications of using a reference
model in feature selection problems, we show its application in the
form of the projection predictive approach, which uses the reference
model to project the posterior draws to the submodels
\citep{paper:original_proj}. The \texttt{R}-package \texttt{projpred}
(available at \url{https://CRAN.R-project.org/package=projpred})
implements the projection in case of submodels in the family of the
generalised linear models and, additionally, it provides a framework
to choose the optimal submodel size through a predictive utility
comparison between the submodels and the reference model.

We now summarise the workflow of the projection predictive approach in
the particular case of the draw-by-draw projection \cite[original
formulation by][]{paper:original_proj}, following
\cite{paper:projpred}. Suppose we have observed $n$ statistical units
with target values $\{y_{i}\}_{i=1}^{n}$ and a set of observed
covariates for which we want to obtain a minimally relevant
subset. Than, the main steps are the following:
\begin{enumerate}
\item Devise and fit a reference model. Let $\{\boldsymbol{\theta}_{*}^{s}\}_{s=1}^{S}$ be the set of $S$ draws from the reference model posterior.
\item Rank the covariates according to their relevance using some
  heuristics and consider as candidate submodels only those which
  preserve this order, starting from including only the highest ranked
  covariate (the submodels are then naturally identified by their
  model size). This is not strictly necessary but reduces the number
  of submodels to consider in the following steps.
\item For each submodel, project the reference model posterior draws as follows:
\begin{equation}\label{eq:draw_by_draw}
\boldsymbol{\theta}_{\perp}^{s} = \underset{\boldsymbol{\theta^{s}}\in\Theta}{\text{argmin}} \frac{1}{n}\sum_{i=1}^{n}\text{KL} \left( p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})\;||\;q(\tilde{y}_{i}|\boldsymbol{\theta}^{s}) \right), \quad s=1,..,S \,,
\end{equation}
where $p(\tilde{y}_{i}|\boldsymbol{\theta}_{*}^{s})$ stands for the
predictive distribution of the reference model with parameters fixed
at $\boldsymbol{\theta}_{*}^{s}$ and conditioning on all the covariate
values related to the statistical unit (identified by the subscript
$i$), whereas $q(\tilde{y}_{i}|\boldsymbol{\theta}^{s})$ is the
predictive distribution of the submodel. The projected draws
$\boldsymbol{\theta}_{\perp}^{s}$ then present the projected posterior
for the submodel.
\item For each submodel (size), test the predictive performance, for example, via cross-validation.
\item Choose the smallest submodel (size) that is sufficiently close to the reference model's predictive utility score.
\end{enumerate}

@@

@@

@@

\hypertarget{shrinkage-signal}{%
\subsubsection{Shrinkage and signal analysis}\label{shrinkage-signal}}
In the framework of fully Bayesian estimation methods, we analyse the
effect of the reference model using two continuous hierarchical
shrinkage priors: the Dirichlet-Laplace
\citep{paper:dirichlet_laplace} and the regularised horseshoe prior
\citep{paper:rhs}.
%
For the Dirichlet hyperparameter of the Dirichlet-Laplace prior, we
use an uniform prior on the interval $(1/p,1/2)$. For the regularised
horseshoe prior, we follow the recommendations provided in
\cite{paper:rhs} by setting the prior guess of the number of relevant
covariates to one in order to achieve the highest possible
shrinkage. The results are not sensitive with respect to the specific
choice.  \ Since, as explained at the start of Section
\ref{comparison}, continuous priors usually do not provide an
automatic selection procedure, we study the amount of shrinkage and
the bias of the posterior medians looking at the SESE and the SSE (see
definitions \eqref{eq:SSE} and \eqref{eq:SESE}), together with the
posterior mean of the shrinkage factors.

@@

@@

@@

@@
